---
title: "homework_ggplot"
author: "Yin Liu"
date: "2023-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Necessary libraries and data input

```{r}
library(readr)
library(ggplot2)
library(tidyverse)
job_postings <- read_csv("updated/job_postings.csv")
```

## 1. Inspect the data for issues. Find anything?

Many columns have a lot of missing values (>90%), so I delete them I keep the columns I am interested in the most (max_salary, min_salary, ...). Also I found that the pay period are different across all the observations. So I change all the payment in yearly, and delete all the annual income that doesn't make sense (<1000 USD/year).

```{r}

## dealing with missings
is.na(job_postings) %>% colSums() ## check for uninformation columns
columns_to_keep <- setdiff(names(job_postings), c("med_salary", "closed_time","skills_desc",
                                                  "application_url","remote_allowed","posting_domain",
                                                  "compensation_type","applies"))
job_postings <- job_postings %>% select(all_of(columns_to_keep))
job <- job_postings[complete.cases(job_postings), ]

## check the pay_period (monthly, hourly, yearly)
job %>% group_by(pay_period) %>% tally() %>% arrange(desc(n)) ## change to yearly
job <- job %>%
  mutate(
    max_salary = case_when(
      pay_period == "HOURLY" ~ max_salary * 40 * 52,
      pay_period == "MONTHLY" ~ max_salary * 12,
      TRUE ~ max_salary
    ),
    min_salary = case_when(
      pay_period == "HOURLY" ~ min_salary * 40 * 52,
      pay_period == "MONTHLY" ~ min_salary * 12,
      TRUE ~ min_salary
    ),
    pay_period = "YEARLY"
  )

job <- job[job$max_salary>=1000,]
```

## 2. Make the following histograms:
  max_salary
  min_salary
  formatted_work_type (Work Type)
  location

```{r}
## histogram of max_salary (yearly), some obeservations do not seem right (<1000)
ggplot(job, aes(x = max_salary)) +
  geom_histogram(bins=200)+  # Adjust binwidth as needed
  labs(title = "Histogram of Max Salary",
       x = "Max Salary",
       y = "Frequency")

ggplot(job, aes(x = min_salary)) +
  geom_histogram(bins=200)+  # Adjust binwidth as needed
  labs(title = "Histogram of Min Salary",
       x = "Min Salary",
       y = "Frequency")

job %>% group_by(formatted_work_type) %>% tally() %>% arrange(desc(n)) 
ggplot(job, aes(x = formatted_work_type, fill = formatted_work_type)) +
  geom_bar() +
  labs(title = "Bar Plot of Work Type",
       x = "Different Types",
       y = "Count")

```

To plot on location, I need to process the data first, because the location are in different formats (eg. "Santa Clarita, CA", "Texas, United States", "United States","New York City Metropolitan Area")

```{r}
## location
job %>% group_by(location) %>% tally() %>% arrange(desc(n))
jobl <- job$location 
addresses <- jobl %>% str_split(", ") %>% lapply(function(inner_list) {
                if (tail(inner_list, 1) == "United States") {
                  inner_list[[1]]
                } else {
                  inner_list[[length(inner_list)]]
                }
              })  %>% unlist() %>% as.matrix()
addresses_unique <- addresses %>% unique()

## use chatgpt to change addresses into the abbreviated state name
state_abbreviations <-  c(
  "USA", "CA", "IL", "NY", "FL", "VA", "CA", "CA", "CA", "MI",
  "MN", "AZ", "TX", "KS", "NM", "ID", "VA", "NV", "MA", "PA",
  "NJ", "DC", "LA", "OH", "FL", "TX", "MD", "MO", "TX", "CO",
  "OR", "NC", "IN", "ND", "CA", "TX", "NY", "WA", "DE", "MT",
  "TN", "MI", "UT", "GA", "CT", "SC", "NY", "NY", "WI", "OH",
  "NJ", "DC", "NV", "OK", "SC", "FL", "IL", "MO", "AR", "MA",
  "CO", "TN", "TN", "KY", "CA", "ND", "IL", "IA", "HI", "NC",
  "MS", "TN", "NY", "NC", "MO", "SC", "ME", "WA", "NE", "AL",
  "CT", "IA", "NC", "NM", "RI", "TX", "OK", "MN", "KY", "CO",
  "GA", "MO", "TX", "AZ", "PA", "AL", "FL", "MI", "AL", "ME",
  "CA", "FL", "OR", "OH", "MN", "AZ", "MI", "OH", "SD", "LA",
  "CA", "PA", "PA", "CO", "TX", "CA", "NH", "DC", "DC", "FL",
  "UT", "UT", "GA", "WY", "NV", "AR", "ND", "SC", "CA", "WV",
  "NH", "DE", "MA", "OR", "MT", "KY", "SD", "OK", "MD", "RI",
  "AZ", "WI", "IN", "NE", "AK", "WV", "KS", "OH", "MA", "PA",
  "TN", "WY", "FL", "CA"
)

# Now, the "location" variable in the "job" dataset contains state abbreviations
index <- match(addresses,addresses_unique)
job$state <- state_abbreviations[index]

ggplot(data = job, aes(x = reorder(state, -table(state)[state]), fill = state)) +
  geom_bar() +
  labs(
    title = "Histogram of State Abbreviations",
    x = "State Abbreviation",
    y = "Frequency"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 4),  # Adjust size as needed
    legend.key.size = unit(0.5, "lines")  # Adjust the legend key size as needed
  )


```


## 3. Now that you've made the histograms what problems have you detected?

The majority of job positions in the data set offer an annual salary ranging from 20,000 to 200,000 USD.
The majority of job positions in the data set are full time.
The majority of job positions in the data set are in CA


## 4. Examine the relationship between max_salary shape and State. Anything interesting?

The max salary of a job offering in CA is higher than the one in the united state overall.

```{r}


ggplot() +
  geom_density(data = job %>% filter(state == "CA"),
               aes(x = max_salary, fill = state), alpha = 0.5, position = "dodge") +
  geom_density(data = job ,
               aes(x = max_salary), color = "blue") +
  labs(title = "Density Plot of Max Salary by of CA and USA overall",
       x = "Max Salary",
       y = "Density") +
  theme_minimal()


```